### 2017-10-31 submission/manuscript.Rmd
complete first result section

#Load in linear mixed-effect data tables
stool_lme <- as.data.frame(read_csv("../data/process/tables/alpha_mixeffect_results.csv"))
unmatch_tissue_lme <- as.data.frame(read_csv("../data/process/tables/alpha_unmatched_mixeffect_results.csv"))
match_tissue_lme <- as.data.frame(read_csv("../data/process/tables/alpha_matched_mixeffect_results.csv"))

@@

***Fecal Diversity is Lower in Those with Carcinoma:***
Using power transformed and Z-score normalized alpha diversity metrics both evenness and the Shannon diversity metrics in feces was lower in those with carcinoma then in controls but not for tissue samples [Figure 1]. Using linear mixed-effects to control for study and variable region there was a significant decrease from control to adenoma to carcinoma for both evenness (P-value = `r paste(format(filter(stool_lme, alpha_metric == "shannoneven")[, "BH"], digits = 2, nsmall = 2))`) and Shannon diversity (P-value = `r paste(format(filter(stool_lme, alpha_metric == "shannon")[, "BH"], digits = 2, nsmall = 2))`). This effect was not observed in tissue when additionally controlling for whether the sample came from the same individual (P-value `r ifelse(min(match_tissue_lme$BH) < 0.05 & min(unmatch_tissue_lme$BH) < 0.05, invisible("check tables"), invisible(">"))` 0.05).

>>>>

#Load in linear mixed-effect alpha diversity data tables 
stool_lme <- as.data.frame(read_csv("../data/process/tables/alpha_mixeffect_results.csv"))
unmatch_tissue_lme <- as.data.frame(read_csv("../data/process/tables/alpha_unmatched_mixeffect_results.csv"))
match_tissue_lme <- as.data.frame(read_csv("../data/process/tables/alpha_matched_mixeffect_results.csv"))

#Load in relative risk alpha diversity data tables
alpha_RR_crc_stool <- as.data.frame(read_csv("../data/process/tables/alpha_RR_composite.csv"))
alpha_RR_adn_stool <- as.data.frame(read_csv("../data/process/tables/alpha_adn_RR_composite.csv"))

alpha_RR_crc_unmatch_tissue <- as.data.frame(read_csv("../data/process/tables/alpha_RR_unmatched_tissue_composite.csv"))
alpha_RR_crc_match_tissue <- as.data.frame(read_csv("../data/process/tables/alpha_RR_matched_tissue_composite.csv"))

alpha_RR_adn_tissue <- as.data.frame(read_csv("../data/process/tables/alpha_adn_RR_tissue_composite.csv"))

@@

***Fecal Diversity is Lower in Those with Carcinoma and increases Relative Risk:***
Using power transformed and Z-score normalized alpha diversity metrics both evenness and the Shannon diversity metrics in feces are lower in those with carcinoma then in controls but not for tissue samples [Figure 1]. Using linear mixed-effects to control for study and variable region there was a significant decrease from control to adenoma to carcinoma for both evenness (P-value = `r paste(format(filter(stool_lme, alpha_metric == "shannoneven")[, "BH"], digits = 2, nsmall = 2))`) and Shannon diversity (P-value = `r paste(format(filter(stool_lme, alpha_metric == "shannon")[, "BH"], digits = 2, nsmall = 2))`). This effect was not observed in tissue when additionally controlling for whether the sample came from the same individual (P-value `r ifelse(min(match_tissue_lme$BH) < 0.05 & min(unmatch_tissue_lme$BH) < 0.05, invisible("check tables"), invisible(">"))` 0.05). For fecal samples a decrease in Shannon diversity and evenness resulted in a significantly increased relative risk for carcinoma (P-value = `r paste(format(filter(alpha_RR_crc_stool, measure == "shannon")[, "pvalue"], digits = 2, nsmall = 2))` and P-value = `r paste(format(filter(alpha_RR_crc_stool, measure == "shannoneven")[, "pvalue"], digits = 2, nsmall = 2))`, respectively) [Figure 2]. Although these values were significant the effect size was relatively small for both metrics (Shannon RR = `r paste(round(filter(alpha_RR_crc_stool, measure == "shannon")[, "rr"], digits = 2))` and evenness RR = `r paste(round(filter(alpha_RR_crc_stool, measure == "shannoneven")[, "rr"], digits = 2))`) [Figure 2]. There was no increased relative risk for these metrics for adenoma or for tissue in general [Figure S1-3]

### 2017-10-31 submission/manuscript.tex
complete first result section

\textbf{\emph{Fecal Diversity is Lower in Those with Carcinoma:}} Using
power transformed and Z-score normalized alpha diversity metrics both
evenness and the Shannon diversity metrics in feces was lower in those
with carcinoma then in controls but not for tissue samples {[}Figure
1{]}. Using linear mixed-effects to control for study and variable
region there was a significant decrease from control to adenoma to
carcinoma for both evenness (P-value = 0.025) and Shannon diversity
(P-value = 0.043). This effect was not observed in tissue when
additionally controlling for whether the sample came from the same
individual (P-value \textgreater{} 0.05).

>>>>

\textbf{\emph{Fecal Diversity is Lower in Those with Carcinoma and
increases Relative Risk:}} Using power transformed and Z-score
normalized alpha diversity metrics both evenness and the Shannon
diversity metrics in feces are lower in those with carcinoma then in
controls but not for tissue samples {[}Figure 1{]}. Using linear
mixed-effects to control for study and variable region there was a
significant decrease from control to adenoma to carcinoma for both
evenness (P-value = 0.025) and Shannon diversity (P-value = 0.043). This
effect was not observed in tissue when additionally controlling for
whether the sample came from the same individual (P-value \textgreater{}
0.05). For fecal samples a decrease in Shannon diversity and evenness
resulted in a significantly increased relative risk for carcinoma
(P-value = 0.01 and P-value = 0.0011, respectively) {[}Figure 2{]}.
Although these values were significant the effect size was relatively
small for both metrics (Shannon RR = 1.31 and evenness RR = 1.34)
{[}Figure 2{]}. There was no increased relative risk for these metrics
for adenoma or for tissue in general {[}Figure S1-3{]}
